---
title: "Business Data Analysis Report"
subtitle: "CSV Import and Data Cleaning"
author: "Aaryn Owen--547 479 873"
date: "`2025-10-08"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
---
```{r regional sales, include=FALSE}
# Create sample regional sales data
regional_sales <- data.frame(
  region = c("North", "South", "East", "West", "North", "South", "East", "West",
             "North", "South", "East", "West", "North", "South", "East", 
             "North", "South","North"),  
  quarter = c(rep(c("Q1", "Q2", "Q3", "Q4"), each = 4), "Q4", "Q4"),
  revenue = c(125000, 98000, 115000, 102000, 
              138000, 105000, 128000, 118000,
              142000, 89000, 135000, 125000,
              155000, 112000, 148000, 165000, 
              120000, 91250),
  customers = c(450, 380, 420, 395, 
                485, 415, 465, 445,
                520, 350, 485, 470,
                575, 425, 545, 590, 435,324)
)
```

# Data Import and Exploration
We reviewed the revenue data for four region over four quarters. There were a total of 8,174 customers across all regions and quarters. The range of revenue values is between $98,000 and $138,000.

*Total Revenue - All Regions and Quarters*
```{r}
aggregate(revenue ~ region, data = regional_sales, sum)
```

*Average revenue per quarter*

```{r Average Revenue, echo=TRUE}
aggregate(revenue ~ quarter, data = regional_sales, sum)
```

*Number of high-performance records (>$140,000)*

```{r}
regional_sales[regional_sales$revenue > 140000, ]
```

```{r, results='asis'}
    library(knitr)
    regional_sales |>
      kable(caption = "Revenue Data")
```

# Data Quality Assessment

Our sales data is evenly distributed across regions and quarters as you can see in  the tables below. This data is encouraging as it shows our business operations are stable and consistent.

```{r}
table(regional_sales$region)
table(regional_sales$quarter)
```

However, our South region shows concerning performance in Q3 with revenue under $90,000. We may want to look further into this region and what factors lead to the lower numbers.

```{r}
regional_sales[regional_sales$revenue < 90000, ]
```


# Business Logic Analysis

The North and East regions are achieving high performance in Q3 and Q4.

```{r}
regional_sales[regional_sales$revenue > 140000, ]
```

We have identified Q4 records with revenue greater than $130,000 and more than 500 customers per region. This combination indicates that we are well-situated for the next fiscal year with strong customer engagement and to achieve our revenue goals for the year.

```{r}
premium_Q4_records <- regional_sales[regional_sales$quarter == "Q4" &
                                       regional_sales$revenue > 130000 &
                                       regional_sales$customers > 500,]
premium_Q4_records[, c("quarter", "revenue", "customers")]
```

The average revenue from the North or East regions is $134,225 which is higher than the combined overall average revenue.

```{r}
records_northeast_regions <- regional_sales[regional_sales$revenue > 0 &
                                  (regional_sales$region == "North" | 
                                    regional_sales$region == "East"), ]
records_northeast_regions[, c("region", "revenue")]
```

Looking at our records that have either revenue below $100,000 or fewer than 400 customers, the revenue below $100,000 represents 0.13% of the total records and fewer than 400 customers represents 34.51% of the total records.

```{r}
concerning_performance <- regional_sales[regional_sales$revenue < 100000 | 
                                           regional_sales$customers < 400, ]
concerning_performance[, c("revenue", "customers")]
```


# Data Aggregation and Business Insights

The North region generated the highest total revenue - $816,250 across all quarters.

```{r}
aggregate(revenue ~ region, data = regional_sales, sum)
```

The highest average customer count was in Q4. Market conditions, consumer behaviour, competitive average, and resource availability could all be business factors that may explain this pattern.

```{r}
aggregate(customers ~ quarter, data = regional_sales, mean)
```


# Executive Summary

Based on our analysis, we have identified the North and East regions as showing the most consistent performance with their strongest results in Q4.

```{r}
records_northeast_regions <- regional_sales[regional_sales$revenue > 0 &
                                  (regional_sales$region == "North" | 
                                    regional_sales$region == "East"), ]
records_northeast_regions[, c("region", "revenue")]
```

```{r}
premium_Q4_records <- regional_sales[regional_sales$quarter == "Q4" &
                                       regional_sales$revenue > 130000 &
                                       regional_sales$customers > 500,]
premium_Q4_records[, c("quarter", "revenue", "customers")]
```